Towards Compiler-Agnostic Performance in Finite-Difference Codes. (TBC)

### 2014-11-11 shallow_perf.tex
Add tables on hardware and compilers

\subsection{The 'shallow' Program}
For this work we have used a benchmarking program that solves the
shallow-water equations on a bi-periodic plane following the
finite-difference scheme introduced by Sadourney.  Originally written
by Paul Schwartrauber of NCAR.  Organic development - optimised for
previous hardware.

@@

\subsection{The {PS}y{KA}l Approach}
Proliferation of lightweight cores is resulting in increased
complexity in the programming models required to achieve performance.
Combine with the complexity of e.g. a model simulating circulation in
the global ocean. Significant challenge in software engineering.
Propose an approach for meeting this challenge by separating the
computational science demands from the e.g. oceanographic demands.

\begin{figure}
\centering
\caption{The {PS}y{KA}l Approach.}
\end{figure}

While encapsulation makes it easier to produce more robust software,
it can often come with a performance penalty. In this work we
investigate the origin of this penalty and the steps required to
remove it.

\section{Results}

We have results for the Gnu, Intel, Cray and IBM compilers.
We've run on Intel Ivy Bridge, Intel Haswell, and IBM BG/Q.

@@

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

@@

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

@@

\section{Conclusion}
The conclusion goes here.

@@

% use section* for acknowledgement
\section*{Acknowledgment}

>>>>

\subsection{The `shallow' Program}
For this work we have used a benchmarking program that solves the
shallow-water equations on a bi-periodic plane following the
finite-difference scheme introduced by Sadourney~\cite{sadourny75}.
Originally written by Paul Schwartrauber of NCAR.  Organic development
- optimised for previous hardware.

In what follows we consider square domains of dimension 64, 128, 256,
512 and 1024.

@@

\subsection{The \psykal Approach}
Proliferation of lightweight cores is resulting in increased
complexity in the programming models required to achieve performance.
Combine with the complexity of e.g. a model simulating circulation in
the global ocean. Significant challenge in software engineering.
Propose an approach for meeting this challenge by separating the
computational science demands from the e.g. oceanographic demands.

\begin{figure}
\centering
\caption{The \psykal Approach.}
\end{figure}

While encapsulation makes it easier to produce more robust software,
it can often come with a performance penalty. In this work we
investigate in detail the hardware and compiler dependencies of this
penalty and the steps required to remove it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology(?)}

Our aim in this work is to achieve portable performance. Consequently,
we have performed tests with the hardware and compilers listed in
Table~\ref{TABLE_compilers}. Where a compiler is available on a given
piece of hardware, the version number used in this work is specified.
The Intel Haswell-based system (Xeon E5-1620 v2) has a clock speed of
3.7~GHz and 10~MB of last-level cache. The Intel Ivy Bridge-based
system (Xeon E5-2697) has a clock speed of 2.7~GHz and a last-level
cache of 30~MB. The CPU in the IBM BG/Q system is built around the
PowerPC A2 core with a clock speed of 1.6~GHz. The chip has 32~MB of
last-level cache.

For completeness, we give the compiler flags used with each compiler
in Table~\ref{TABLE_compiler_flags}. These flags are particularly
important for the \psykal versions of the code. The performance of the
original version of the code is very much less sensitive to the
compiler options. This is because it consists of a single source file
and a single routine (excluding those for I/O).

We do retain I/O because any realistic code has to output results to disk.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{The compilers and CPUs used in this work.}
\label{TABLE_compilers}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
                 & Gnu   & Intel       & Cray    & IBM bgxlf\\
\hline
Intel Haswell    & 4.8.2 & 14.0.0      &         &          \\
Intel Ivy Bridge &       & 14.0.1.106  & 8.3.3   &          \\
IBM A2 (BGQ)     &       &             &         & 14.1     \\
IBM Power 8      &       &             &         &          \\
\hline
\end{tabular}
\end{table}

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{The compiler flags used in this work.}
\label{TABLE_compiler_flags}
\centering
\begin{tabular}{|l|c|}
\hline
Compiler  &  Flags \\
\hline
Gnu       & -Ofast -mtune=native -finline-limit=50000    \\
Intel     & -O3 -fast -fno-inline-factor    \\
Cray      & -O3 -O ipa5 -h wp    \\
IBM bgxlf & -O5 -qinline=auto:level=10    \\
\hline
\end{tabular}
\end{table}

@@

Beginning with the vanilla \psykal version, we apply a series of
optimisations with the aim of recovering, as much as is possible, the
structure and thus performance of the original version of the code.

\begin{enumerate}

\item Pass explicit array bounds to middle layer (rather than using
  assumed-size array arguments)

\item Move kernel routines into same module as middle layer

\item Loop fusion

\item In-lined field-copy operation

\item Fused field-copy operation

\end{enumerate}

These optimisations do not always result in improved performance (AND
HENCE WE SHOULD CALL THEM TRANSFORMATIONS PERHAPS?). Whether or not
they do so depends both on the compiler and the problem size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

In figure~\ref{FIG_orig_perf_summary} we plot the performance of the
original version of the shallow code for the range of compilers and
hardware considered here. This summary demonstrates the effect of the
larger last-level cache of the Intel Ivy Bridge system compared to
that of the Haswell system; note the significant drop in performance
in going from a problem size of $256^{2}$ to a problem size of
$512^{2}$ for the black and yellow bars. The performance of the Ivy
Bridge system (red and purple bars) only drops to this level when the
domain size is increased to $1024^{2}$. At this point, the working set
no longer fits within cache and performance is dominated by the
bandwidth to main memory making it relatively insensitive to the
choice of compiler.

For problem sizes that fit within cache there is
little to choose between the performance obtained from the Intel and
Gnu compilers.

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{orig_summary}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
\caption{Summary of the performance of the original shallow code on 
the range of compilers and hardware under consideration.}
\label{FIG_orig_perf_summary}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{best_psykal_summary}
\caption{Summary of the best performance achieved by any \psykal 
version of shallow with each of the compilers and CPUs under 
consideration.}
\label{FIG_best_psykal_perf_summary}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{slowdown_summary}
\caption{Summary of the performance achieved by any \psykal
version as a percentage of that achieved by the original version of the code.}
\label{FIG_slowdown_summary}
\end{figure}

@@

@@

\section{Conclusions}
The conclusion goes here.

@@

% use section* for acknowledgement
\section*{Acknowledgments}
